SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-08 docs/questioning.tex
some changes - more on the discussion

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} records all database
transactions without tagging each transaction with the
corresponding call-site. In this case a false negative result
is prominent. For instance, consider an application that will send the
password for a forgetful user via email by executing the following
query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit}, could be easily escaped
since a malicious user, could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\cite{APKLM10}.
%% PL Like the return-to-lib attack?

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking (see
Subsubsection~\ref{sec:taint}) together with parse-tree validation.
Diglossia was based on the theory of Ray and Ligatti~\cite{RL12b}
(also see Section~\ref{sec:attacks}) to detect {\sc dsl} code
injection attacks. In addition, it is actually the first, and so far
the only framework that detects No{\sc sql} injection attacks. When an
application computes an output string (query), Diglossia computes a
``shadow" of this string. Specifically, it maps all characters
introduced by the application to a shadow character set. This set does
not contain any characters coming from the tainted input. Then, the
scheme creates the tree of the query that is about to be executed and
compares it with the parsed tree of the ``shadow" (locus {\sc z}). If
the trees are not equal the application is probably under attack. Note
that Diglossia can be bypassed in the same way as other taint
tracking approaches as described in reference~\cite{NBR14}.
%% PL which way is this exactly?

@@

Table~\ref{tab:comp} indicates that the authors of 3 out of 34
({\bf 8.8\%}) publications, provided complete quadruples.
On the other hand, we see that 4 out of
34 ({\bf 11.7\%}) were not tested at all.
In this cases, all the four elements of the
quadruples are not available ({\sc na}).
% Maybe we should comment on the fact that these are
% frameworks coming from the etiological category
% and detect the attacks ``by design".
An interesting observation is that 10 out of 34
({\bf 29.41\%}) of the authors performed only attacks
during their tests and did not test their mechanisms
to see if they produce any false alarms. The corresponding
quadruples in this case, are the ones that contain {\sc tp} and
{\sc fn} results but {\sc tn} and {\sc fp} results are not
available ({\sc na}). This stresses the fact that authors
maybe more interested into making sure that
their mechanisms can detect known attacks than seeing how
they respond under normal conditions.
In some occasions (where the number of {\sc tp} results
are not quantified), authors did not mention how many or
which attacks they performed.
Also, in some cases, we observe that even if
the authors report the existence of possible
false positive and negative results, they have not
quantified them ({\sc nq}).

@@

There are also authors that took a different approach.
For instance, during their initial tests,
Stock et al.~\cite{SLMS14} managed to bypass
the browser-based {\sc xss} filters of the
73\% of 1,602 real-world {\sc dom}-based {\sc xss} vulnerabilities.
These vulnerabilities were actually found during
their previous research~\cite{LSJ13}.
Furthermore, in the publications that
{\sc amnesia}~\cite{HO05,HO06,HO05b}
and {\sc sd}river~\cite{MS09,MKS09,MKLS11} were presented,
authors managed to break existing application suites
and then test the diagnostic performance of their tools on them.

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 ({\bf 23\%}) do so at the level
of the {\sc rdbms} --- locus $\Omega$.
The other 11 wrap {\sc api} calls related to output
vectors --- locus {\sc z}.

However, in the case of frameworks that deal
with JavaScript-driven {\sc xss} attacks
we observe that attack detection may take place
either on the server-side or on the client-side.
In the first case, a proxy is placed on the server
to examine the server's responses before they
reach the user's web browser. In the second
case a modified browser or a library that
is securely downloaded from the server,
checks the responses for potential attacks.
Our findings indicate that the tendency so far,
is to create frameworks that perform detection
on the client side. In particular, 11 out of
19 frameworks ({\bf 57,8\%}) detect JavaScript-driven
{\sc xss} attacks at the client's side.
Note that the majority of the mechanisms that
detect such attacks on the server-side, can also
detect {\sc dsl} code injection attacks (7 out of 8).

@@

Our first observation indicates that only 6
out of 34 ({\bf 17.6\%}) of the authors provided a homepage for
their mechanism through their publication.
2 from these 6 homepages are currently not available.
Also, in 5 cases, authors made either the source code
or their executables available after their paper
got published. Finally, we observe that only 3
out of 34 ({\bf 8.8\%}) authors have currently
their testbeds available on-line.

@@

>>>>

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly}
do not tag transactions with their corresponding call-sites.
In this case a false negative result
is prominent. For instance, consider an application that will send the
password for a forgetful user via email by executing the following
query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit} and {\it libAnomaly}, could be easily escaped
since a malicious user, could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\cite{APKLM10}.
%% PL Like the return-to-lib attack?

@@

{\it Diglossia}~\cite{SMS13} combines positive taint tracking (see
Subsubsection~\ref{sec:taint}) together with parse-tree validation.
Diglossia was based on the theory of Ray and Ligatti~\cite{RL12b}
(also see Section~\ref{sec:attacks}) to detect {\sc dsl} code
injection attacks. In addition, it is actually the first, and so far
the only framework that detects No{\sc sql} injection attacks. When an
application computes an output string (query), Diglossia computes a
``shadow" of this string. Specifically, it maps all characters
introduced by the application to a shadow character set. This set does
not contain any characters coming from the tainted input. Then, the
scheme creates the tree of the query that is about to be executed and
compares it with the parsed tree of the ``shadow" (Locus {\sc z}). If
the trees are not equal the application is probably under attack. Note
that Diglossia can be bypassed in the same way as other taint
tracking approaches as described in reference~\cite{NBR14}.
%% PL which way is this exactly?

@@

Table~\ref{tab:comp} indicates that the authors of 3 out of 34
(8.8\%) publications, provided complete quadruples.
On the other hand, we see that 4 out of
34 (11.7\%) were not tested at all.
In this cases, all the four elements of the
quadruples are not available ({\sc na}).
% Maybe we should comment on the fact that these are
% frameworks coming from the etiological category
% and detect the attacks ``by design".
An interesting observation is that 10 out of 34
(29.41\%) of the authors performed only attacks
during their tests and did not test their mechanisms
to see if they produce any false alarms. The corresponding
quadruples in this case, are the ones that contain {\sc tp} and
{\sc fn} results but {\sc tn} and {\sc fp} results are not
available ({\sc na}). This stresses the fact that authors
maybe more interested into making sure that
their mechanisms can detect known attacks than seeing how
they respond under normal conditions.
In some occasions (where the number of {\sc tp} results
are not quantified), authors did not mention how many or
which attacks they performed.
Also, in some cases, we observe that even if
the authors report the existence of possible
false positive and negative results, they have not
quantified them ({\sc nq}).

@@

There are also authors that took a different approach.
For instance, during their initial tests,
Stock et al.~\cite{SLMS14} managed to bypass
the browser-based {\sc xss} filters of the
73\% of 1,602 real-world {\sc dom}-based {\sc xss} vulnerabilities.
These vulnerabilities were actually found during
their previous research~\cite{LSJ13}.
Finally, in the publications that
{\sc amnesia}~\cite{HO05,HO06,HO05b}
and {\sc sd}river~\cite{MS09,MKS09,MKLS11} were presented,
authors managed to break existing application suites
and then test the diagnostic performance of their tools on them.

@@

Given the fact that the final steps of
{\sc dsl} injection and {\sc csrf} attacks take place
on the server-side, all the mechanisms that detect
such attacks are placed in some point within the server.
From the mechanisms that detect {\sc dsl} code injection
attacks, 3 out of 13 (23\%) do so at the level
of the {\sc rdbms} --- locus $\Omega$.
The other 11 wrap {\sc api} calls related to output
vectors --- locus {\sc z}.

However, in the case of frameworks that deal
with JavaScript-driven {\sc xss} attacks
we observe that attack detection may take place
either on the server-side or on the client-side.
In the first case, a proxy is placed on the server
to examine the server's responses before they
reach the user's web browser. In the second
case a modified browser or a library that
is securely downloaded from the server,
checks the responses for potential attacks.
Our findings indicate that the tendency so far,
is to create frameworks that perform detection
on the client side. In particular, 11 out of
19 frameworks (57,8\%) detect JavaScript-driven
{\sc xss} attacks at the client's side.
Note that the majority of the mechanisms that
detect such attacks on the server-side, can also
detect {\sc dsl} code injection attacks (7 out of 8).

@@

Our first observation indicates that only 6
out of 34 (17.6\%) of the authors provided a homepage for
their mechanism through their publication.
2 from these 6 homepages are currently not available.
Also, in 5 cases, authors made either the source code
or their executables available after their paper
got published. Finally, we observe that only 3
out of 34 (8.8\%) authors have currently
their testbeds available on-line.

@@

We believe that our research may facilitate researchers
or practitioners who would like to use one of the
aforementioned mechanisms and researchers who would like
to build new ones. Especially in the second case,
researchers can utilize our exploitation model as a
starting point to develop their approach.
